---
title: "TJPW Metrics"
author: "Lester Blumberg"
date: "12/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
```

# Version 1: Singles Matches Only from 2013-14

```{r load-data}
tjpw13 = read_excel("TJPW Metrics.xlsx", sheet = 2, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 14)))

tjpw14 = read_excel("TJPW Metrics.xlsx", sheet = 3, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 14)))

tjpw15 = read_excel("TJPW Metrics.xlsx", sheet = 4, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 14)))

tjpw16 = read_excel("TJPW Metrics.xlsx", sheet = 5, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 14)))

tjpw = rbind(tjpw13, tjpw14, tjpw15, tjpw16)
```

Subset by match type
```{r subset-by-match}
singles = tjpw %>% 
  filter(`Match Type` == "Singles")

tag = tjpw %>% 
  filter(`Match Type` == "Tag")

threeway = tjpw %>% 
  filter(`Match Type` == "3 Way")

sixperson = tjpw %>% 
  filter(`Match Type` == "6 Person Tag")
```

```{r}
summary(factor(tjpw$Event))
```

Formula for Expected Outcome
```{r expected-outcome-function}
expected <- function(ra, rb, s = 400){
  expected_outcome = 1 / (1 + 10^((rb - ra)/s))
  return(expected_outcome)
} 
```

Formula for New Rating
```{r new-rating-function}
new_rating <- function(ra, rb, outcome, k = 20){
  new = ra + k * (outcome - expected(ra, rb))
  return(new)
}

new_rating_threeway <- function(r1, r2, r3, outcome, k = 20){
  expected_threeway = (expected(r1, r2) + expected(r1, r3)) / 3
  new = r1 + k * (outcome - expected_threeway)
  return(new)
}
```


```{r dates-events}

events_dates = tjpw %>% 
  select(Event, Date) %>% 
  distinct(Event, .keep_all = T)
```

```{r wrestlers}
wrestlers = as.vector(rbind(tjpw$`Wrestler 1`, tjpw$`Wrestler 2`, tjpw$`Wrestler 3`,
                            tjpw$`Wrestler 4`, tjpw$`Wrestler 5`, tjpw$`Wrestler 6`)) %>% 
  unique() %>% 
  na.omit()
```

Create new dataframe for ratings with events, dates, and each wrestler as columns
```{r ratings-dataframe}
Ratings = data.frame(matrix(nrow = nrow(events_dates), ncol = length(wrestlers)+2))
colnames(Ratings) = c("Event", "Date", wrestlers)
Ratings$Event = events_dates$Event
Ratings$Date = events_dates$Date

starting_ratings = c(1500, 1500, rep(NA, length(wrestlers)-2))
```

Function for Calculating the new ratings for an individual event
```{r calculation-function}
#V1
calculate_new_ratings <- function(indiv_event){
  this_event = singles %>% 
    filter(Event == indiv_event)
  
  prior_ratings = if(indiv_event == "The First Meeting"){
    starting_ratings
  }
  else{
    Ratings[which(Ratings$Event == indiv_event) - 1,3:ncol(Ratings)]
  }
  
  new_ratings = prior_ratings
  for(i in 1:nrow(this_event)){
    W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
    W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
    prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500)
    prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500)
    if(this_event$Decision[i] == "Time Limit Draw"){
      new_ratings[W1-2] =  new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 0.5)
      new_ratings[W2-2] =  new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0.5)
    }
    else{
      new_ratings[W1-2] =  new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 1)
      new_ratings[W2-2] =  new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0)
    }
  }
  return(new_ratings)
}
```

Function for Calculating Ratings for an Individual Event (currently includes functionality for singles, tag, 6 person tag, and 3 way matches)
```{r}
#V2.3.2 (Include 2016 Data)
calculate_new_ratings2 <- function(indiv_event){
  this_event = tjpw %>% 
    filter(Event == indiv_event)
  
  prior_ratings = if(indiv_event == "The First Meeting"){
    starting_ratings
  }
  else{
    Ratings[which(Ratings$Event == indiv_event) - 1,3:ncol(Ratings)]
  }
  
  new_ratings = prior_ratings
  for(i in 1:nrow(this_event)){
    if(this_event$`Match Type`[i] == "Singles"){
       W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
       W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
       prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2],
                                    1500)
       prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2],
                                    1500)
       if(this_event$Decision[i] == "Time Limit Draw"){
         new_ratings[W1-2] =  new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 0.5)
         new_ratings[W2-2] =  new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0.5)
       }
       else{
         new_ratings[W1-2] =  new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 1)
         new_ratings[W2-2] =  new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0)
       }
    }
    if(this_event$`Match Type`[i] == "Tag"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      W4 = which(colnames(Ratings) == this_event$`Wrestler 4`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500)
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500)
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500)
      prior_ratings[W4-2] = ifelse(!is.na(prior_ratings[W4-2]), prior_ratings[W4-2], 1500)
      prior_team1_rating = mean(c(as.double(prior_ratings[W1-2]),
                                  as.double(prior_ratings[W2-2])))
      prior_team2_rating = mean(c(as.double(prior_ratings[W3-2]),
                                  as.double(prior_ratings[W4-2])))
      if(this_event$Decision[i] == "Time Limit Draw"){
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 0.5)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0.5)
      }
      else{
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
      }
      new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (2/3)
      new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/3)
      new_ratings[W3-2] = prior_ratings[W3-2] + (new_team2_rating - prior_team2_rating) * (2/3)
      new_ratings[W4-2] = prior_ratings[W4-2] + (new_team2_rating - prior_team2_rating) * (1/3)
    }
    if(this_event$`Match Type`[i] == "6 Person Tag"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      W4 = which(colnames(Ratings) == this_event$`Wrestler 4`[i])
      W5 = which(colnames(Ratings) == this_event$`Wrestler 5`[i])
      W6 = which(colnames(Ratings) == this_event$`Wrestler 6`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500)
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500)
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500)
      prior_ratings[W4-2] = ifelse(!is.na(prior_ratings[W4-2]), prior_ratings[W4-2], 1500)
      prior_ratings[W5-2] = ifelse(!is.na(prior_ratings[W5-2]), prior_ratings[W5-2], 1500)
      prior_ratings[W6-2] = ifelse(!is.na(prior_ratings[W6-2]), prior_ratings[W6-2], 1500)
      prior_team1_rating = mean(c(as.double(prior_ratings[W1-2]),
                                  as.double(prior_ratings[W2-2]),
                                  as.double(prior_ratings[W3-2])))
      prior_team2_rating = mean(c(as.double(prior_ratings[W4-2]),
                                  as.double(prior_ratings[W5-2]),
                                  as.double(prior_ratings[W6-2])))
      if(this_event$Decision[i] == "Time Limit Draw"){
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 0.5)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0.5)
      }
      else{
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
      }
      new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (1/2)
      new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/4)
      new_ratings[W3-2] = prior_ratings[W3-2] + (new_team1_rating - prior_team1_rating) * (1/4)
      new_ratings[W4-2] = prior_ratings[W4-2] + (new_team2_rating - prior_team2_rating) * (1/2)
      new_ratings[W5-2] = prior_ratings[W5-2] + (new_team2_rating - prior_team2_rating) * (1/4)
      new_ratings[W6-2] = prior_ratings[W6-2] + (new_team2_rating - prior_team2_rating) * (1/4)
    }
    if(this_event$`Match Type`[i] == "3 Way"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500)
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500)
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500)
      if(this_event$Decision[i] == "Time Limit Draw"){
        new_ratings[W1-2] = new_rating_threeway(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 1/3)
        new_ratings[W2-2] = new_rating_threeway(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 1/3)
        new_ratings[W3-2] = new_rating_threeway(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 1/3)
      }
      else{
        if(this_event$`W3 Result`[i] == "L*"){
          if(this_event$`W2 Result`[i] == "L*"){
            new_ratings[W1-2] = new_rating_threeway(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 1)
            new_ratings[W2-2] = new_rating_threeway(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 0)
            new_ratings[W3-2] = new_rating_threeway(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 0)
          }
          if(this_event$`W2 Result`[i] == "W*"){
            new_ratings[W1-2] = new_rating_threeway(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 0.5)
            new_ratings[W2-2] = new_rating_threeway(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 0.5)
            new_ratings[W3-2] = new_rating_threeway(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 0)
          }
        }
        else{
          new_ratings[W1-2] = new_rating_threeway(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 5/6)
          new_ratings[W2-2] = new_rating_threeway(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 0)
          new_ratings[W3-2] = new_rating_threeway(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 1/6)
        }
      }
    }
    prior_ratings = new_ratings #Fixes issue with multiple matches at one show
  }
  return(new_ratings)
}
```


Apply functions to dataset to get ratings from start to finish
```{r run-ratings}
j = 1
while(j <= nrow(Ratings)){
  Ratings[j, 3:ncol(Ratings)] = calculate_new_ratings2(events_dates$Event[j])
  j = j+1
}
```

Make data longform for easier plotting
```{r longer}
Ratings.long = Ratings %>% 
  pivot_longer(3:ncol(Ratings), names_to = "Wrestler", values_to = "Rating")

long.2016 = Ratings %>% 
  select(c(1:4,6,8:9,11,13:16,18:24,26:27)) %>% 
  pivot_longer(`Miyu Yamashita`:`Mil Clown`, names_to = "Wrestler", values_to = "Rating")
```

```{r plot}
Ratings.long %>% 
  ggplot(aes(x = Date, y = Rating, color = Wrestler)) +
  geom_line(size = 1) +
  scale_color_viridis_d() +
  geom_hline(yintercept = 1500) +
  labs(title = "TJPW Elo Ratings v2.3.2 2013-2015",
       subtitle = "s=400, k=20, Starting Value 1500")

long.2016 %>% 
  ggplot(aes(x = Date, y = Rating, color = Wrestler)) +
  geom_line(size = 1) +
  scale_color_viridis_d() +
  geom_hline(yintercept = 1500) +
  labs(title = "TJPW Elo Ratings v2.3.2 2013-2016 (only active in 2016)",
       subtitle = "s=400, k=20, Starting Value 1500")
```

```{r}
end2013 = Ratings[7, 3:7]
row.names(end2013) = "Elo Rating"

end2014 = Ratings[28, 3:14]
row.names(end2014) = "Elo Rating"

end2015 = Ratings[46, 3:21]
row.names(end2015) = "Elo Rating"

end2016 = Ratings[79, 3:27]
row.names(end2016) = "Elo Rating"

Ratings.v2.3.2 = Ratings
```

```{r}
elo.change.2013 = end2013 - rep(1500, 5)
row.names(elo.change.2013) = "Rating Change"

elo.change.2014 = end2014 - c(end2013, rep(1500, 7))
row.names(elo.change.2014) = "Rating Change"
  
elo.change.2015 = end2015 - c(end2014, rep(1500,7))
row.names(elo.change.2015) = "Rating Change"
  
elo.change.2016 = end2016 - c(end2015, rep(1500,6))
row.names(elo.change.2016) = "Rating Change"
```


# Including Tag Matches

I'm not sure the best way to approach this, but it feels like a wrestler should not gain or lose as many points for a tag match as for a singles match. One approach would be to see that the total points gained or lost for a team is roughly equivalent to the points gained or lost for a wrestler in a singles match.

The other questions are how to determine how much each member of a team gains and loses and which wrestler from the opposing team they gain and lose from. Should the team member who picks up the win get 75% of team points? 2/3? Should the team member who is defeated lose 75% of team points? 2/3?

# Including 3-Way Matches

This is another difficult consideration. Clearly, the winning wrestler deserves to gain points, and the wrestler who is defeated deserves to lose points. However, how should I consider the wrestler who is not involved in the decision? Also, how should the points transaction compare to a 1v1 singles match? While a 3-way victory is impressive, the accomplishment is not the same as defeating a wrestler 1 on 1.

# Other match types

Decisions made for tag and 3-way matches should be instructive and consistent with how I include match types not seen yet, such as multi-person tag matches (3v3, 4v4, etc), free-for-all matches with more than 3 competitors (4-way, battle royal, etc), and 3-way and 4-way tag matches.

# Titles and Tournaments

2014 sees the first edition of the Tokyo Princess Cup, a single elimination tournament featuring 8 wrestlers. Should each match be treated purely as a singles match? Should there be any kind of bonus for reaching a certain round or winning the whole thing? Similarly, when titles enter the picture, should title matches be treated as ordinary matches? Elo ratings are constructed to give and take points equally. Would I have to take extra points away from the loser of a tournament or title match to give those points to the winner? Would that be the right call?

# Matches outside of TJPW

TJPW wrestlers competed in dark matches at DDT Peter Pan in both 2013 and 2014. These matches are theoretically no different from matches that would take place as part of a TJPW show. Should I include them in the ratings? Should this exception only be made for TJPW offer matches at DDT shows or also for offer matches at other promotions' shows (if they exist)?

# Handling irregular and outside wrestlers

Without ratings for the larger joshi wrestling scene, it will be difficult to give proper ratings to outsiders who come into TJPW for special matches every so often, such as Aja Kong or Meiko Satomura. Saki Akai is one such wrestler already in the dataset, with only 2 matches in TJPW. With enough matches, any new wrestler should be fairly quickly moved in the right direction in the ratings, but there might not be enough matches for all outsiders. Should special ratings be given to debuting outside wrestlers whose body of work makes it clear that they are far beyond a rookie TJPW wrestler?

# Match Duration

Match duration is anecdotally linked to wrestler parity. Early on, many of the matches are short because all of the wrestlers are fairly new to the sport. However, there will be larger disparities in wrestler experience in the future. It only follows that a wrestler in their first year will be defeated by a strong wrestler more quickly than they would if they faced off two years later. Should the duration of a match play a role in the ratings change for the match?

# Post 2015 Data Thoughts

## Wrestlers with new names

Ririko Kendo became Rika Tatsumi. This is not a gimmick change. Just a new name. I'll need to devise a way to have them count as the same person, possibly just manually. Also, Saki Akai will be difficult. I'll have to see if Saki Akamiya counts enough as an entirely different person to just make separate, but I know that Sakisama is coming. It seems like that's just a continuation of Biishiki-Gun Saki Akai. This will take some consideration.

## Tag Points

Tag teams must be included. The ratings are missing too much right now. I think just pool the wrestlers on each team for average team ratings and treat the match like a singles match between two wrestlers with those ratings. Then just divvy up for who defeated whom.

## Tag Bonus

It might be a good idea to add a tag bonus to the prior rating for teams that have teamed up before. While there are not a lot of designated teams, they do exist, and it does seem like they perform better than just a random assortment of wrestlers. Two ideas for a bonus: either based on # of previous matches (total or in a given period of time) or based on prior win percentage as a team.

## Battle Royal

Still no clue how to factor that in. My best thought is to give rating points to the winner, give a small amount to the runner up, and take small but equal amounts from every other participant to make this happen. In the end, I think that winning a battle royal should be worth more than just winning an individual match. Maybe this will be reflected through expected outcome.

## 3 way

Still unsure about this too. Maybe play with expected outcome (starting from 1/3 instead of 1/2) like battle royal idea.

# V2 Complete

## Tag Team Functionality

This is a good starting point. Currently, the member of the team who scores the victory gets 2/3 of the credit. Similarly, the defeated member gets hit with 2/3 of the penalty.

## Six Person Tag

Probably just similar to 2v2 tag. Maybe start with a breakdown of 1/2, 1/4, 1/4 for the three team members with the half going to the person getting the win/being defeated.

## Using the ratings to answer some questions

Once I have enough data being fed in, I should be able to use my ratings to answer some of my questions for fine tuning, such as tag bonus and how to treat outside wrestlers.

# 12/21/21 Notes

I have realized that my function to calculate ratings for an individual event have a flaw for when a wrestler has more than one match at an event. While this is a fairly rare occurrence, it does happen a few times, especially for the semifinals and finals of the Princess Cup. I believe this should be a simple fix where I have the prior ratings update at the end of the for loop for each individual match. This should also enable me to undo the Ai Shimizu Appearance patch and go back to treating that as one event. If I ever want to check what the elo ratings would be in the middle of a show, it shouldn't be that hard to figure out manually.

## Post-Implementation

The change seems to have worked and was very simple. Now, I'll try to remove the Ai Shimizu Appearance patch and see if the results are the same.

## Post Patch Removal

Took it out. It's all still good. Next step is to add in trios matches.

## Post Trios

Pretty simple inclusion. Currently 1/2 credit to victor and defeated. 1/4 to other team participants.

## Post Threeway

Well, the implementation works. Might need changes later if more types of 3 way finishes occur. Currently, the expected outcome is derived by taking the sum of expected outcomes against each opponent and dividing by 3. This sums to 1 for the three competitors and seems to make sense. For the actual outcome, in a typical match with one winner, one defeated participant, and one loser not involved in the decision, I am currently giving 2/3 as the outcome to the winner, 0 to the main loser, and 1/3 to the uninvolved party. Other considerations are 3/4 to the winner and 1/4 to uninvolved as it feels like there should at least be a slight penalty for not winning. Another idea is to try to give the winner the same boost in a completely even match that they would get in a singles match, which requires the actual outcome being 0.5 greater than the expected outcome. So this would set the actual outcomes at 5/6 for the winner and 1/6 for the loser. This approach has the effect of making the gain equal for the two match types when the competitors are closer in prior rating but gives more credit to the winner of a 3 way match than the winner of a singles match when the winner has a much higher prior rating than the other competitors.

This will take more thought in the future and will probably be easier to determine when there is more data.

As of right now, the only matches not in the system are a battle royal and a handicap match. I do not wish to program ratings for these until I have more of each.

## V2.3.1

I flipped and decided to make the outcomes 5/6 and 1/6 for 3 way matches. After looking at it a bit, I think that this helps to make sure that the winners get around the same boost as winning a singles match. Also, the uninvolved loser still takes some penalty, which is fair since the goal of wrestling is to win the match. Also, the penalty is still about half of what it would be if they were the defeated loser, so they are still shielded some. I will probably leave this alone for now and focus more on adding 2016 into the dataset.

# 12/22/21 Notes

I finished data entry for 2016 last night. No new Battle Royals this year. There is one match of a new type though. There was a threeway elimination match. I might wait before coding its inclusion just to see if there are more matches like this in 2017 or 2018. 2016 also saw a few instances of multiple shows on the same day. I fixed an issue with creating the Ratings dataframe that should fix any possible issues there. I also removed the one match from a DDT show that I had been including. Looking at Cagematch shows that there are a number of these matches that occur over the years. For now, it is best to exclude them. I can always go back and add them in later. One exception will be if any important matches (such as a title match) take place on a DDT show.

## Titles

2016 sees the introduction of the (at this point in time) TOKYO Princess Of Princess Title. Yamashita holds it for most of the year, but impressive rookie Yuu wins it off of her. Because I'm currently using an ELO rating, I don't think I should give any special significance to title matches. However, I am noting them in the spreadsheet. If I move on to a more customized metric, I will definitely look at including title reigns, title matches, and #1 contender matches.

## Mil Clown

Late in 2016, Yuka Sakazaki departs to Magic Country while her twin sister Mil Clown arrives in her place. Mil Clown has a better performance record than her sister. It is tempting to combine their matches, but they are two different people. I will keep them separate for now.

## Post 2016 Data Inclusion

Really interesting stuff. The greater number of events and matches leads to some big changes. Yuu gets the Samoa Joe build and gains 125 on top of the 1510 she ended 2015 with after her debut. Yuka and Shoko also see big jumps this year. Yuka isn't quite back to 1500 yet, but if I were to include her sister's matches as hers, she would definitely get there. Shoko's big year sees her become the 3rd wrestler behind Miyu and Yuu to cross the 1600 barrier. Once I include the elimination match, she'll only end the year even higher. There are also big losses, even bigger than the big gains. Azusa Takigawa ends 2016 in hot pursuit of a husband; for her sake, I hope she's found the bottom of the Elo Ratings. She dropped a whopping 166 points in the ratings. She finishes the year with the worst rating seen so far, and she has gone over a year into her career having never seen her rating improve once from a match. Nonoko teaming with her is a true sign of friendship. Rookies Nodoka and Marika have also seen their initial ratings fall over 100 points in 2016. This will probably become common occurrence for rookies since they will be wrestling a large number of matches and have a roster full of wrestlers more experienced than them.

One interesting note is that Miyu has only posted small gains of 12 and 25 in '15 and '16 respectively. I wonder if this will become a trend that my current rating system has somewhat of an upper bound or if it will be easier for Miyu to climb higher when more wrestlers get over 1600.

We're starting to see some veterans like Hyper Misao and Rika win more matches. 2017 might be the year we see them get above and stay above 1500.

## Elimination Match

If there isn't another one in 2017, I'll program it in. Since the result was Shoko pinning both Akane and Miyu, Shoko will definitely get full credit for the win. I feel that an elimination match should count for more than one standard match, singles or threeway. Perhaps the best way to treat it is as two matches. A threeway first and then a singles match between the remaining competitors. If I were to do that, I would want to credit differently for the first part than 5/6 and 1/6 since the remaining competitor should not be penalized for not being involved in the first elimination. Part of an elimination match is just sticking around as long as you can. It makes sense to reward Shoko for the first pinfall because any time you pin an opponent is an example of your wrestling prowess. Another option is to treat it as two singles matches, but I feel that this might understate the likelihood of losing the match during the first stage. This will take some thought. It might be good to peruse the remaining years to see if there are any more of these matches. If there are, then I can code it in a way to be flexible for future match inclusion.