---
title: "TJPW Metrics"
author: "Lester Blumberg"
date: "12/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
```

# Version 1: Singles Matches Only from 2013-14

```{r load-data}
tjpw13 = read_excel("TJPW Metrics.xlsx", sheet = 2, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 19)))

tjpw14 = read_excel("TJPW Metrics.xlsx", sheet = 3, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 19)))

tjpw15 = read_excel("TJPW Metrics.xlsx", sheet = 4, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 19)))

tjpw16 = read_excel("TJPW Metrics.xlsx", sheet = 5, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 19)))

tjpw17 = read_excel("TJPW Metrics.xlsx", sheet = 6, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 19)))

tjpw18 = read_excel("TJPW Metrics.xlsx", sheet = 7, col_types = c("guess", "date", "guess", "guess", "guess", "text", rep("guess", 19)))

tjpw = rbind(tjpw13, tjpw14, tjpw15, tjpw16, tjpw17, tjpw18)
```

Subset by match type
```{r subset-by-match}
singles = tjpw %>% 
  filter(`Match Type` == "Singles")

tag = tjpw %>% 
  filter(`Match Type` == "Tag")

threeway = tjpw %>% 
  filter(`Match Type` == "3 Way")

sixperson = tjpw %>% 
  filter(`Match Type` == "6 Person Tag")

eightperson = tjpw %>% 
  filter(`Match Type` == "8 Person Tag")

handicap = tjpw %>% 
  filter(`Match Type` %in% c("1 v 2 Handicap", "2 v 3 Handicap"))

elimination = tjpw %>% 
  filter(`Match Type` == "Elimination")

battle_royal = tjpw %>% 
  filter(`Match Type` == "Battle Royal")
```

```{r}
summary(factor(tjpw$Event))
```

Formula for Expected Outcome
```{r expected-outcome-function}
expected <- function(ra, rb, s = 400){
  expected_outcome = 1 / (1 + 10^((rb - ra)/s))
  return(expected_outcome)
} 

expected_threeway <- function(ra, rb, rc){
  expected_outcome_threeway = (expected(ra, rb) + expected(ra, rc)) / 3
  return(expected_outcome_threeway)
}

expected_elim3 <- function(ra, rb, rc){
  expected_outcome_elim3 = 1 - ((expected(rb, ra) + expected(rc, ra)) / 3)
  return(expected_outcome_elim3)
}
```

Formula for New Rating
```{r new-rating-function}
new_rating <- function(r1, r2, outcome, k = 20){
  new = r1 + k * (outcome - expected(r1, r2))
  return(new)
}

new_rating_threeway <- function(r1, r2, r3, outcome, k = 20){
  new = r1 + k * (outcome - expected_threeway(r1, r2, r3))
  return(new)
}

new_rating_elim3 <- function(r1, r2, r3, outcome, k = 20){
  new = r1 + k * (outcome - expected_elim3(r1, r2, r3))
  return(new)
}
```


```{r dates-events}

events_dates = tjpw %>% 
  select(Event, Date) %>% 
  distinct(Event, .keep_all = T)
```

```{r wrestlers}
wrestlers = as.vector(rbind(tjpw$`Wrestler 1`, tjpw$`Wrestler 2`, tjpw$`Wrestler 3`,
                            tjpw$`Wrestler 4`, tjpw$`Wrestler 5`, tjpw$`Wrestler 6`,
                            tjpw$`Wrestler 7`, tjpw$`Wrestler 8`)) %>% 
  unique() %>% 
  na.omit()
```

```{r prior-experience}
prior_exp = read_excel("TJPW Metrics.xlsx", sheet = 8)
```

```{r draws}
# to include all types of draws that exist in database
draws = c("Time Limit Draw", "Double Countout")
```

```{r wins-and-losses-sets}
wins = c("W", "W*")
losses = c("L", "L*")
```

Create new dataframe for ratings with events, dates, and each wrestler as columns
```{r ratings-dataframe}
Ratings = data.frame(matrix(nrow = nrow(events_dates), ncol = length(wrestlers)+2))
colnames(Ratings) = c("Event", "Date", wrestlers)
Ratings$Event = events_dates$Event
Ratings$Date = events_dates$Date

starting_ratings = c(1500, 1500, rep(NA, length(wrestlers)-2))
```

Function for Calculating the new ratings for an individual event (1st draft)
```{r calculation-function}
#V1
calculate_new_ratings <- function(indiv_event){
  this_event = singles %>% 
    filter(Event == indiv_event)
  
  prior_ratings = if(indiv_event == "The First Meeting"){
    starting_ratings
  }
  else{
    Ratings[which(Ratings$Event == indiv_event) - 1,3:ncol(Ratings)]
  }
  
  new_ratings = prior_ratings
  for(i in 1:nrow(this_event)){
    W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
    W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
    prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500)
    prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500)
    if(this_event$Decision[i] == "Time Limit Draw"){
      new_ratings[W1-2] =  new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 0.5)
      new_ratings[W2-2] =  new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0.5)
    }
    else{
      new_ratings[W1-2] =  new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 1)
      new_ratings[W2-2] =  new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0)
    }
  }
  return(new_ratings)
}
```

Function for Calculating Ratings for an Individual Event (evolving with the times)
```{r current-calculation-function}
#V2.4.3 (Added elimination singles matches)
calculate_new_ratings2 <- function(indiv_event){
  this_event = tjpw %>% 
    filter(Event == indiv_event)
  
  prior_ratings = if(indiv_event == "The First Meeting"){
    starting_ratings
  }
  else{
    Ratings[which(Ratings$Event == indiv_event) - 1,3:ncol(Ratings)]
  }
  
  new_ratings = prior_ratings
  for(i in 1:nrow(this_event)){
    if(this_event$`Match Type`[i] == "Singles"){
       W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
       W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
       prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2],
                                    1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 1`[i])])
       prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2],
                                    1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 2`[i])])
       if(this_event$Decision[i] %in% draws){
         new_ratings[W1-2] =  new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 0.5)
         new_ratings[W2-2] =  new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0.5)
       }
       if(this_event$Decision[i] == "No Contest"){
         new_ratings[W1-2] = prior_ratings[W1-2]
         new_ratings[W2-2] = prior_ratings[W2-2]
       }
       else{
         new_ratings[W1-2] =  new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 1)
         new_ratings[W2-2] =  new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0)
       }
    }
    if(this_event$`Match Type`[i] == "Tag"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      W4 = which(colnames(Ratings) == this_event$`Wrestler 4`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 1`[i])])
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 2`[i])])
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 3`[i])])
      prior_ratings[W4-2] = ifelse(!is.na(prior_ratings[W4-2]), prior_ratings[W4-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 4`[i])])
      prior_team1_rating = mean(c(as.double(prior_ratings[W1-2]),
                                  as.double(prior_ratings[W2-2])))
      prior_team2_rating = mean(c(as.double(prior_ratings[W3-2]),
                                  as.double(prior_ratings[W4-2])))
      if(this_event$Decision[i] %in% draws){
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 0.5)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0.5)
        new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * 0.5
        new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * 0.5
        new_ratings[W3-2] = prior_ratings[W3-2] + (new_team2_rating - prior_team2_rating) * 0.5
        new_ratings[W4-2] = prior_ratings[W4-2] + (new_team2_rating - prior_team2_rating) * 0.5
      }
      if(this_event$Decision[i] == "No Contest"){
        new_ratings[W1-2] = prior_ratings[W1-2]
        new_ratings[W2-2] = prior_ratings[W2-2]
        new_ratings[W3-2] = prior_ratings[W3-2]
        new_ratings[W4-2] = prior_ratings[W4-2]
      }
      else{
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
        new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (2/3)
        new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/3)
        new_ratings[W3-2] = prior_ratings[W3-2] + (new_team2_rating - prior_team2_rating) * (2/3)
        new_ratings[W4-2] = prior_ratings[W4-2] + (new_team2_rating - prior_team2_rating) * (1/3)
      }
    }
    if(this_event$`Match Type`[i] == "6 Person Tag"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      W4 = which(colnames(Ratings) == this_event$`Wrestler 4`[i])
      W5 = which(colnames(Ratings) == this_event$`Wrestler 5`[i])
      W6 = which(colnames(Ratings) == this_event$`Wrestler 6`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 1`[i])])
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 2`[i])])
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 3`[i])])
      prior_ratings[W4-2] = ifelse(!is.na(prior_ratings[W4-2]), prior_ratings[W4-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 4`[i])])
      prior_ratings[W5-2] = ifelse(!is.na(prior_ratings[W5-2]), prior_ratings[W5-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 5`[i])])
      prior_ratings[W6-2] = ifelse(!is.na(prior_ratings[W6-2]), prior_ratings[W6-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 6`[i])])
      prior_team1_rating = mean(c(as.double(prior_ratings[W1-2]),
                                  as.double(prior_ratings[W2-2]),
                                  as.double(prior_ratings[W3-2])))
      prior_team2_rating = mean(c(as.double(prior_ratings[W4-2]),
                                  as.double(prior_ratings[W5-2]),
                                  as.double(prior_ratings[W6-2])))
      if(this_event$Decision[i] %in% draws){
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 0.5)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0.5)
        new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (1/3)
        new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/3)
        new_ratings[W3-2] = prior_ratings[W3-2] + (new_team1_rating - prior_team1_rating) * (1/3)
        new_ratings[W4-2] = prior_ratings[W4-2] + (new_team2_rating - prior_team2_rating) * (1/3)
        new_ratings[W5-2] = prior_ratings[W5-2] + (new_team2_rating - prior_team2_rating) * (1/3)
        new_ratings[W6-2] = prior_ratings[W6-2] + (new_team2_rating - prior_team2_rating) * (1/3)
      }
      if(this_event$Decision[i] == "No Contest"){
        new_ratings[W1-2] = prior_ratings[W1-2]
        new_ratings[W2-2] = prior_ratings[W2-2]
        new_ratings[W3-2] = prior_ratings[W3-2]
        new_ratings[W4-2] = prior_ratings[W4-2]
        new_ratings[W5-2] = prior_ratings[W5-2]
        new_ratings[W6-2] = prior_ratings[W6-2]
      }
      else{
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
        new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (1/2)
        new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/4)
        new_ratings[W3-2] = prior_ratings[W3-2] + (new_team1_rating - prior_team1_rating) * (1/4)
        new_ratings[W4-2] = prior_ratings[W4-2] + (new_team2_rating - prior_team2_rating) * (1/2)
        new_ratings[W5-2] = prior_ratings[W5-2] + (new_team2_rating - prior_team2_rating) * (1/4)
        new_ratings[W6-2] = prior_ratings[W6-2] + (new_team2_rating - prior_team2_rating) * (1/4)
      }
    }
    if(this_event$`Match Type`[i] == "8 Person Tag"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      W4 = which(colnames(Ratings) == this_event$`Wrestler 4`[i])
      W5 = which(colnames(Ratings) == this_event$`Wrestler 5`[i])
      W6 = which(colnames(Ratings) == this_event$`Wrestler 6`[i])
      W7 = which(colnames(Ratings) == this_event$`Wrestler 7`[i])
      W8 = which(colnames(Ratings) == this_event$`Wrestler 8`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 1`[i])])
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 2`[i])])
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 3`[i])])
      prior_ratings[W4-2] = ifelse(!is.na(prior_ratings[W4-2]), prior_ratings[W4-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 4`[i])])
      prior_ratings[W5-2] = ifelse(!is.na(prior_ratings[W5-2]), prior_ratings[W5-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 5`[i])])
      prior_ratings[W6-2] = ifelse(!is.na(prior_ratings[W6-2]), prior_ratings[W6-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 6`[i])])
      prior_ratings[W7-2] = ifelse(!is.na(prior_ratings[W7-2]), prior_ratings[W7-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 7`[i])])
      prior_ratings[W8-2] = ifelse(!is.na(prior_ratings[W8-2]), prior_ratings[W8-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 8`[i])])
      prior_team1_rating = mean(c(as.double(prior_ratings[W1-2]),
                                  as.double(prior_ratings[W2-2]),
                                  as.double(prior_ratings[W3-2]),
                                  as.double(prior_ratings[W4-2])))
      prior_team2_rating = mean(c(as.double(prior_ratings[W5-2]),
                                  as.double(prior_ratings[W6-2]),
                                  as.double(prior_ratings[W7-2]),
                                  as.double(prior_ratings[W8-2])))
      if(this_event$Decision[i] %in% draws){
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 0.5)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0.5)
        new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (1/4)
        new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/4)
        new_ratings[W3-2] = prior_ratings[W3-2] + (new_team1_rating - prior_team1_rating) * (1/4)
        new_ratings[W4-2] = prior_ratings[W4-2] + (new_team1_rating - prior_team1_rating) * (1/4)
        new_ratings[W5-2] = prior_ratings[W5-2] + (new_team2_rating - prior_team2_rating) * (1/4)
        new_ratings[W6-2] = prior_ratings[W6-2] + (new_team2_rating - prior_team2_rating) * (1/4)
        new_ratings[W7-2] = prior_ratings[W7-2] + (new_team2_rating - prior_team2_rating) * (1/4)
        new_ratings[W8-2] = prior_ratings[W8-2] + (new_team2_rating - prior_team2_rating) * (1/4)
      }
      if(this_event$Decision[i] == "No Contest"){
        new_ratings[W1-2] = prior_ratings[W1-2]
        new_ratings[W2-2] = prior_ratings[W2-2]
        new_ratings[W3-2] = prior_ratings[W3-2]
        new_ratings[W4-2] = prior_ratings[W4-2]
        new_ratings[W5-2] = prior_ratings[W5-2]
        new_ratings[W6-2] = prior_ratings[W6-2]
        new_ratings[W7-2] = prior_ratings[W7-2]
        new_ratings[W8-2] = prior_ratings[W8-2]
      }
      else{
        new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
        new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
        new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (2/5)
        new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/5)
        new_ratings[W3-2] = prior_ratings[W3-2] + (new_team1_rating - prior_team1_rating) * (1/5)
        new_ratings[W4-2] = prior_ratings[W4-2] + (new_team1_rating - prior_team1_rating) * (1/5)
        new_ratings[W5-2] = prior_ratings[W5-2] + (new_team2_rating - prior_team2_rating) * (2/5)
        new_ratings[W6-2] = prior_ratings[W6-2] + (new_team2_rating - prior_team2_rating) * (1/5)
        new_ratings[W7-2] = prior_ratings[W7-2] + (new_team2_rating - prior_team2_rating) * (1/5)
        new_ratings[W8-2] = prior_ratings[W8-2] + (new_team2_rating - prior_team2_rating) * (1/5)
      }
    }
    if(this_event$`Match Type`[i] == "3 Way"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 1`[i])])
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 2`[i])])
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 3`[i])])
      if(this_event$Decision[i] %in% draws){
        new_ratings[W1-2] = new_rating_threeway(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 1/3)
        new_ratings[W2-2] = new_rating_threeway(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 1/3)
        new_ratings[W3-2] = new_rating_threeway(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 1/3)
      }
      if(this_event$Decision[i] == "No Contest"){
        new_ratings[W1-2] = prior_ratings[W1-2]
        new_ratings[W2-2] = prior_ratings[W2-2]
        new_ratings[W3-2] = prior_ratings[W3-2]
      }
      else{
        if(this_event$`W3 Result`[i] == "L*"){
          if(this_event$`W2 Result`[i] == "L*"){
            new_ratings[W1-2] = new_rating_threeway(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 1)
            new_ratings[W2-2] = new_rating_threeway(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 0)
            new_ratings[W3-2] = new_rating_threeway(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 0)
          }
          if(this_event$`W2 Result`[i] == "W*"){
            new_ratings[W1-2] = new_rating_threeway(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 0.5)
            new_ratings[W2-2] = new_rating_threeway(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 0.5)
            new_ratings[W3-2] = new_rating_threeway(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 0)
          }
        }
        else{
          new_ratings[W1-2] = new_rating_threeway(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 5/6)
          new_ratings[W2-2] = new_rating_threeway(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 0)
          new_ratings[W3-2] = new_rating_threeway(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 1/6)
        }
      }
    }
    if(this_event$`Match Type`[i] == "1 v 2 Handicap"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 1`[i])])
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 2`[i])])
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 3`[i])])
      if(this_event$Decision[i] %in% draws){
# Left Blank to be filled in if necessary        
      }
      if(this_event$Decision[i] == "No Contest"){
        new_ratings[W1-2] = prior_ratings[W1-2]
        new_ratings[W2-2] = prior_ratings[W2-2]
        new_ratings[W3-2] = prior_ratings[W3-2]
      }
      else{
        if(this_event$`W1 Result`[i] == "W"){#indicates W1 is alone
          prior_team1_rating = as.double(prior_ratings[W1-2]) / 2
          prior_team2_rating = mean(c(as.double(prior_ratings[W2-2]),
                                      as.double(prior_ratings[W3-2])))
          new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
          new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
          if(this_event$`W3 Result`[i] == "L"){
            new_ratings[W1-2] = prior_ratings[W1-2] + new_team1_rating - prior_team1_rating
            new_ratings[W2-2] = prior_ratings[W2-2] + (new_team2_rating - prior_team2_rating) * (2/3)
            new_ratings[W3-2] = prior_ratings[W3-2] + (new_team2_rating - prior_team2_rating) * (1/3)
          }
          #Can add on if result occurs where both W2 and W3 are defeated simultaneously
        }
        else{#indicates W1 is part of the team in the handicap match
          prior_team1_rating = mean(c(as.double(prior_ratings[W1-2]),
                                      as.double(prior_ratings[W2-2])))
          prior_team2_rating = as.double(prior_ratings[W3-2]) / 2
          new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
          new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
          if(this_event$`W2 Result`[i] == "W"){
            new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (2/3)
            new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/3)
            new_ratings[W3-2] = prior_ratings[W3-2] + new_team2_rating - prior_team2_rating
          }
          #Can add on if result occurs where both W1 and W2 defeat W3 simultaneously
        }
      }
    }
    if(this_event$`Match Type`[i] == "2 v 3 Handicap"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      W4 = which(colnames(Ratings) == this_event$`Wrestler 4`[i])
      W5 = which(colnames(Ratings) == this_event$`Wrestler 5`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 1`[i])])
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 2`[i])])
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 3`[i])])
      prior_ratings[W4-2] = ifelse(!is.na(prior_ratings[W4-2]), prior_ratings[W4-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 4`[i])])
      prior_ratings[W5-2] = ifelse(!is.na(prior_ratings[W5-2]), prior_ratings[W5-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 5`[i])])
      if(this_event$Decision[i] %in% draws){
# Left Blank to be filled in if necessary        
      }
      if(this_event$Decision[i] == "No Contest"){
        new_ratings[W1-2] = prior_ratings[W1-2]
        new_ratings[W2-2] = prior_ratings[W2-2]
        new_ratings[W3-2] = prior_ratings[W3-2]
      }
      else{
        if(this_event$`W3 Result`[i] %in% losses){#indicates 2-person team won
          prior_team1_rating = mean(c(as.double(prior_ratings[W1-2]),
                                      as.double(prior_ratings[W2-2])))
          prior_team2_rating = mean(c(as.double(prior_ratings[W3-2]),
                                      as.double(prior_ratings[W4-2]),
                                      as.double(prior_ratings[W5-2])))
          new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
          new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
          if(this_event$`W2 Result`[i] == "W" & this_event$`W4 Result`[i] == "L"){
            new_ratings[W1-2] = prior_ratings[W1-2] + (new_team1_rating - prior_team1_rating) * (2/3)
            new_ratings[W2-2] = prior_ratings[W2-2] + (new_team1_rating - prior_team1_rating) * (1/3)
            new_ratings[W3-2] = prior_ratings[W3-2] + (new_team2_rating - prior_team2_rating) * (1/2)
            new_ratings[W4-2] = prior_ratings[W4-2] + (new_team2_rating - prior_team2_rating) * (1/4)
            new_ratings[W5-2] = prior_ratings[W5-2] + (new_team2_rating - prior_team2_rating) * (1/4)
          }
          #Can add on later if results occur where multiple wrestlers on Team 2 are defeated simultaneously and/or both wrestlers on Team 1 win simultaneously
        }
        else{#indicates 3-person team won
          prior_team1_rating = mean(c(as.double(prior_ratings[W1-2]),
                                      as.double(prior_ratings[W2-2]),
                                      as.double(prior_ratings[W3-2])))
          prior_team2_rating = mean(c(as.double(prior_ratings[W4-2]),
                                      as.double(prior_ratings[W5-2])))
          new_team1_rating = new_rating(prior_team1_rating, prior_team2_rating, 1)
          new_team2_rating = new_rating(prior_team2_rating, prior_team1_rating, 0)
          if(this_event$`W2 Result`[i] == "W" & this_event$`W5 Result`[i] == "L"){
            new_ratings[W1-2] = prior_ratings[W1-2] + (new_team2_rating - prior_team2_rating) * (1/2)
            new_ratings[W2-2] = prior_ratings[W2-2] + (new_team2_rating - prior_team2_rating) * (1/4)
            new_ratings[W3-2] = prior_ratings[W3-2] + (new_team2_rating - prior_team2_rating) * (1/4)
            new_ratings[W4-2] = prior_ratings[W4-2] + (new_team1_rating - prior_team1_rating) * (2/3)
            new_ratings[W5-2] = prior_ratings[W5-2] + (new_team1_rating - prior_team1_rating) * (1/3)
          }
          #Can add on later if results occur where multiple wrestlers on Team 1 win simultaneously and/or both wrestlers on Team 2 are defeated simultaneously
        }
      }
    }
    if(this_event$`Match Type`[i] == "Elimination"){
      W1 = which(colnames(Ratings) == this_event$`Wrestler 1`[i])
      W2 = which(colnames(Ratings) == this_event$`Wrestler 2`[i])
      W3 = which(colnames(Ratings) == this_event$`Wrestler 3`[i])
      prior_ratings[W1-2] = ifelse(!is.na(prior_ratings[W1-2]), prior_ratings[W1-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 1`[i])])
      prior_ratings[W2-2] = ifelse(!is.na(prior_ratings[W2-2]), prior_ratings[W2-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 2`[i])])
      prior_ratings[W3-2] = ifelse(!is.na(prior_ratings[W3-2]), prior_ratings[W3-2], 1500 + 20 * prior_exp$Years[which(prior_exp$Wrestler == this_event$`Wrestler 3`[i])])
      if(is.na(this_event$`Wrestler 4`[i])){
        if(this_event$`W1 Result`[i] == "W*W"){
          w1.part1 = new_rating_elim3(prior_ratings[W1-2], prior_ratings[W2-2], prior_ratings[W3-2], 1 + ((1 - expected_elim3(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2])) / 2)) - prior_ratings[W1-2]
          new_ratings[W1-2] = new_rating(prior_ratings[W1-2], prior_ratings[W2-2], 1) + w1.part1
          w2.part1 = new_rating_elim3(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2], 1 - ((1 - expected_elim3(prior_ratings[W2-2], prior_ratings[W1-2], prior_ratings[W3-2])) / 2)) - prior_ratings[W2-2]
          new_ratings[W2-2] = new_rating(prior_ratings[W2-2], prior_ratings[W1-2], 0) + w2.part1
          new_ratings[W3-2] = new_rating_elim3(prior_ratings[W3-2], prior_ratings[W2-2], prior_ratings[W1-2], 0)
        }
      }
    }
    prior_ratings = new_ratings #Fixes issue with multiple matches at one show
  }
  return(new_ratings)
}
```


Apply functions to dataset to get ratings from start to finish
```{r run-ratings}
j = 1
while(j <= nrow(Ratings)){
  Ratings[j, 3:ncol(Ratings)] = calculate_new_ratings2(events_dates$Event[j])
  j = j+1
}
```

Make data longform for easier plotting
```{r longer}
Ratings.long = Ratings %>% 
  pivot_longer(3:ncol(Ratings), names_to = "Wrestler", values_to = "Rating")

long.2017 = Ratings %>% 
  select(c(1:3,6,8:9,11,13,16,18,20:22,24,26:27,30:34,36)) %>% 
  pivot_longer(`Miyu Yamashita`:`Yuki Kamifuku`, names_to = "Wrestler", values_to = "Rating")
```

```{r plot}
Ratings.long %>% 
  ggplot(aes(x = Date, y = Rating, color = Wrestler)) +
  geom_line(size = 1) +
  scale_color_viridis_d() +
  geom_hline(yintercept = 1500) +
  labs(title = "TJPW Elo Ratings v2.4.2 2013-2017",
       subtitle = "s=400, k=20, Starting Value 1500")

long.2017 %>% 
  ggplot(aes(x = Date, y = Rating, color = Wrestler)) +
  geom_line(size = 1) +
  scale_color_viridis_d() +
  geom_hline(yintercept = 1500) +
  labs(title = "TJPW Elo Ratings v2.4.2 2013-2017 (only active in 2017)",
       subtitle = "s=400, k=20, Starting Value 1500")
```

```{r year-end}
end2013 = Ratings[7, 3:10]
row.names(end2013) = "Elo Rating"

end2014 = Ratings[28, 3:15]
row.names(end2014) = "Elo Rating"

end2015 = Ratings[46, 3:21]
row.names(end2015) = "Elo Rating"

end2016 = Ratings[79, 3:27]
row.names(end2016) = "Elo Rating"

end2017 = Ratings[111, 3:39]
row.names(end2017) = "Elo Rating"

Ratings.v2.4.3 = Ratings
write.table(Ratings.v2.4.3, "TJPW Elo Ratings v2.4.3.csv")
```

```{r yearly-changes}
elo.change.2013 = end2013 - rep(1500, 8)
row.names(elo.change.2013) = "Rating Change"

elo.change.2014 = end2014 - c(end2013, rep(1500, 5))
row.names(elo.change.2014) = "Rating Change"
  
elo.change.2015 = end2015 - c(end2014, rep(1500, 6))
row.names(elo.change.2015) = "Rating Change"
  
elo.change.2016 = end2016 - c(end2015, rep(1500, 6))
row.names(elo.change.2016) = "Rating Change"

elo.change.2017 = end2017 - c(end2016, rep(1500,12))
row.names(elo.change.2017) = "Rating Change"

yearend.df = Ratings[c(7, 28, 46, 79, 111), 3:39]
row.names(yearend.df) = c("2013", "2014", "2015", "2016", "2017")
elo.change.df = rbind(yearend.df[1,] - rep(1500, 37),
                      yearend.df[2,] - c(yearend.df[1, 1:8], rep(1500, 29)),
                      yearend.df[3,] - c(yearend.df[2, 1:13], rep(1500, 24)),
                      yearend.df[4,] - c(yearend.df[3, 1:19], rep(1500, 18)),
                      yearend.df[5,] - c(yearend.df[4, 1:25], rep(1500, 12)))
```


